---
title: "GEOtop Calibration through Optimization (hydroPSO)"
author: "Emanuele Cordano"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GEOtop Calibration through Optimization (hydroPSO)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

BLA BLA Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Introduction

Why this work ....
Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## GEOtop Hydrological Model


GEOtop (Endrizzi et al, 2014 and references therein) is a distributed model of the mass and energy balance of the hydrological cycle. GEOtop is applicable to simulations in continuum in small or relatively large montain catchments. GEOtop deals with the effects of topography on the interaction between energy balance and hydrological cycle (water, glacier and snow) with peculiar solutions. The source code of GEOtop 2.0 with detailed documentation is available for researchers and environmental/software engineers through the following links:

* https://code.google.com/p/geotop/
* https://github.com/se27xx/GEOtop/
* https://github.com/skyglobe/geotop

For  Unix-like OS users, a C source code can be rapidly downloaded and built with the following instrunctions (https://github.com/se27xx/GEOtop/):


1. Open a console and go to the drectory where to clone GEOtop source code;
2. Clone the source code typing: "git clone https://github.com/se27xx/GEOtop";
3. Enter GEOtop directory typing: "cd GEOtop";
4. Create the subdirectory for the executable file  "mkdir bin";
5. Build and create GEOtop executale file: "make -f geotop.make"

GEOtop executable will be created in the subdirectory "bin". 
For other versions of GEOtop, please see the related URLs. 

## Overview of the main function

GEOtopOptim2 is a package which is able to enter a GEOtop simulation folder and calibarate some parameters (set by user) optimizing a Goodnes-of-Fit measure between point results and point observations. 


Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.


## Exercise

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.




## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
