---
title: "Verification of a GEOtop point simulation in B2 LTER Site - Match - South Tyrol - Italy"
author: "Johannes Brenner, Emanuele Cordano"
date: "June 2016"
output: html_document
runtime: shiny
---

***

This R Markdown document is made interactive using Shiny. To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

***

```{r, include=FALSE}

library(geotopbricks)
if(!require("geotopbricks"))
{
  if(!require("devtools"))
  {
    install.packages(devtools)
    require("devtools")
  }
  install_github("ecor/geotopbricks")
  require("geotopbricks")
}


if(!require("geotopAnalytics"))
{
  if(!require("devtools"))
  {
    install.packages(devtools)
    require("devtools")
  }
  install_github("ecor/geotopAnalytics")
  require("geotopAnalytics")
}

if(!require("dygraphs"))
{
  install.packages(dygraphs)
  require("dygraphs")
}

if(!require("hydroGOF"))
{
  install.packages(hydroGOF)
  require("hydroGOF")
}

if(!require("ggplot2"))
{
  install.packages(ggplot2)
  require("ggplot2")
}


######wpath <- '/home/ecor/temp/geotopOptim_tests/B2_BeG_017_DVM_001' 
####wpath <- '/home/ecor/temp/geotopOptim_tests/B2_BeG_017_DVM_001_test_1' 
####'/home/ecor/activity/2016/eurac2016/Incarico_EURAC/Simulations/B2/B2_BeG_017_DVM_001_test_2' 
#### TO CHECK ...
##wpath <- '/home/ecor/activity/2016/eurac2016/Incarico_EURAC/results/B2_BeG_017_DVM_001_test_1' 

#wpath <- "/run/user/1000/gvfs/smb-share:server=sdcalp01.eurac.edu,share=data2/Simulations/Simulation_GEOtop_1_225_ZH/Vinschgau/SimTraining/BrJ/HiResAlp/1D/Montecini_pnt_1_225_B2_007/"
#wpath <- "/run/user/1000/gvfs/smb-share:server=sdcalp01.eurac.edu,share=data2/Simulations/Simulation_GEOtop_1_225_ZH/Vinschgau/SimTraining/BrJ/HiResAlp/1D/Montecini_pnt_1_225_P2_003/"
#wpath <- "/run/user/1000/gvfs/smb-share:server=sdcalp01.eurac.edu,share=data2/Simulations/Simulation_GEOtop_1_225_ZH/Vinschgau/SimTraining/BrJ/MonaLisa/1D/Kaltern/sim006/"

# obs data
#load(file.path(wpath, "obs", "observation.RData"))
#names(observation) <- c("hour", "day")
#obs <- observation
#time(obs$hour) <- as.POSIXlt(time(obs$hour))

# sim data  ### WORK HERE 
#if (file.exists(file.path(wpath,"PointOutValidation.RData"))) {
#  load(file.path(wpath,"PointOutValidation.RData"))
#} else {

#}

#varout <- var_out

# adjust datetime: observations -1 hour
#time(obs$hour) <- as.POSIXlt(time(obs$hour))
#time(obs$hour)$mday <- time(obs$hour)$mday - 1
#time(obs$day) <- time(obs$day) - 1

# get hourly lag between obs and sim
## ccf_temp <- stats::ccf(var_out[["air_temperature"]], obs$hour[,"air_temperature"], plot = F, na.action = na.pass)
## lag <- round(ccf_temp$acf[which.max(ccf_temp$acf)],0)
## time(obs$hour)$hour <- time(obs$hour)$hour + lag

# load lookup table
##data("lookup_tbl_observation")

wpath <-  '/home/ecor/activity/2016/eurac2016/idra/B2_BeG_017_DVM_001_test_1' 
alldata <- GEOtop_ReadValidationData(wpath = wpath, save_rData = T)

obsnames <- attr(alldata,"observation_var")
simnames <- attr(alldata,"simulation_var")

fun_names <- c("min","mean","max","sum")
```

#### Interactive Inputs

```{r, echo=FALSE}
inputPanel(
  ##textInput(inputId = "simFolder", label = "Simulation folder", value = ""),
  
  selectInput(inputId = "variable", label = "discover variable", choices = obsnames,selected=obsnames[8]),
  
  selectInput(inputId = "add_variable", label = "additional variable (sim)", choices = simnames, selected = simnames[2]),
  
  radioButtons(inputId = "aggregation", label = "aggregation", choices = c("hourly","daily","monthly"), selected = "hourly", inline = FALSE),
  
  selectInput(inputId = "aggregation_function_var", label = "Aggregation Function", choices = fun_names, selected = "mean"),
  
  selectInput(inputId = "aggregation_function_add_var", label = "Aggregation Function (add_var)", choices = fun_names, selected = "sum")
  
  
  
  
#  radioButtons(inputId = "flux_amount", label = "flux or amount", choices = c("flux","amount"), selected = "flux", inline = FALSE)
  
#  # radioButtons(inputId = "cum", label = "cumulated or time series", choices = c("time series","cumulated over time"), selected = "time series", inline = FALSE)
  
)

```

***

#### Time Series Plot

Dynamic plots of observed vs simulated time series. 

```{r, echo=FALSE}


renderDygraph({
  
  
  ## TO GO ON .....
  	iiv <- input$variable
  	aiiv <- input$add_variable
  	aggregation <- input$aggregation
  	aggr_fun_1 <- input$aggregation_function_var
  	aggr_fun_2 <-  input$aggregation_function_add_var
    data <- extractGeotopVar(x=alldata,InputVar=iiv,Add_InputVar=aiiv,aggregate=aggregation,aggregate_fun=c(aggr_fun_1,aggr_fun_2)) 
    unit <- attr(alldata,"var_unit")[iiv]
    add_unit <- attr(alldata,"var_unit")[aiiv]
    ## 
    
   dygraph(data, ylab=paste(iiv,"[",unit,"]",sep="")) %>%
		dyRangeSelector() %>%
		dyRoller() %>%
		dySeries(name = "additional.var", axis = "y2", stepPlot = TRUE, fillGraph = TRUE, label = paste(aiiv,"[",add_unit,"]",sep="")) %>%
		dyAxis(name="y2",label=paste("[",add_unit,"]",sep=""))




#  }
 
})

```
  
***

####Summary Table on Goodness of Fit (GOF)

Measures for GOF are given for seasons and for the whole data series. Calculations were performed with the hydroGOF R-Package. Type ?gof in the R console for information on specific GOFs.

```{r, echo=FALSE}

renderDataTable({

#renderText({
# if (input$aggregation=="hour" | input$aggregation=="day")
#  {
#    observation <- obs[[input$aggregation]][,input$variable]
#  } else {
#  	observation <- obs[["day"]][,input$variable]
#  }
#  
#  simulation <- varout[[input$variable]]
#  
# 	if (input$aggregation=="day" & input$flux_amount=="flux") simulation <- aggregate(simulation, as.Date(time(simulation)), mean, na.rm=T) 
#  	if (input$aggregation=="day" & input$flux_amount=="amount") simulation <- aggregate(simulation, as.Date(time(simulation)), sum, na.rm=F)
#  	if (input$aggregation=="month" & input$flux_amount=="flux") 
#  {
#    simulation <- aggregate(simulation, as.yearmon(time(simulation)), mean, na.rm=T)
#    observation <- aggregate(observation, as.yearmon(time(observation)), mean, na.rm=T)
#  } 
#  if (input$aggregation=="month" & input$flux_amount=="amount") 
#  {
#    simulation <- aggregate(simulation, as.yearmon(time(simulation)), sum, na.rm=F)
#    observation <- aggregate(observation, as.yearmon(time(observation)), sum, na.rm=F)
#  } 
  
#  time(observation) <- as.POSIXct(time(observation))
#  time(simulation) <- as.POSIXct(time(simulation))
  
#  data <- merge(observation, simulation)
 
 iiv <- input$variable
    aiiv <- input$add_variable
  	nnn <- c("simulation","observation")
  	
  	aggregation <- input$aggregation
  	aggr_fun_1 <- input$aggregation_function_var
  	aggr_fun_2 <-  input$aggregation_function_add_var
    data <- extractGeotopVar(x=alldata,InputVar=iiv,Add_InputVar=aiiv,aggregate=aggregation,aggregate_fun=c(aggr_fun_1,aggr_fun_2)) 
    
  	data <- data[,nnn]
 
  
  
  
  data <- extractGeotopVar(x=alldata,InputVar=iiv,Add_InputVar=aiiv)[,nnn]
  
  gofs <- gofg(sim = data$simulation, obs=data$observation, na.rm = TRUE)
  gofs <- as.data.frame(gofs)
  names(gofs) <- "YEAR"
  gofs$GOF <- dimnames(gofs)[[1]]
  
  mon <- as.numeric(as.character(index(data),format="%m")) ####as.numeric(format(time(data), "%m"))
  
  datadjf <-  data[mon==12 | mon==1 | mon==2,]
  gofs$DJF <-  c(gofg(sim = datadjf$simulation, obs=datadjf$observation, na.rm = TRUE))
  datamam <-  data[mon==3 | mon==4 | mon==5,]
  gofs$MAM <-  c(gofg(sim = datamam$simulation, obs=datamam$observation, na.rm = TRUE))
  datajja <-  data[mon==6 | mon==7 | mon==8,]
  gofs$JJA <-  c(gofg(sim = datajja$simulation, obs=datajja$observation, na.rm = TRUE))
  datason <-  data[mon==9 | mon==10 | mon==11,]
  gofs$SON <-  c(gofg(sim = datason$simulation, obs=datason$observation, na.rm = TRUE))
  
  gofs <- gofs[,c(2,3,4,5,6,1)]

}, options = list(pageLength=5, lengthMenu=c(5, 10, 15, 20)))

```

***

####Scatterplot

```{r, echo=FALSE}

renderPlot({
  
	iiv <- input$variable
    aiiv <- input$add_variable
  	nnn <- c("simulation","observation")
  	
  	aggregation <- input$aggregation
  	aggr_fun_1 <- input$aggregation_function_var
  	aggr_fun_2 <-  input$aggregation_function_add_var
    data <- extractGeotopVar(x=alldata,InputVar=iiv,Add_InputVar=aiiv,aggregate=aggregation,aggregate_fun=c(aggr_fun_1,aggr_fun_2)) 
    
  	data <- data[,nnn]
	
	if (all(is.na(data$observation))==TRUE) {
	
				data$observation <- -9999
	}
	if (all(is.na(data$simulation))==TRUE) {
	
				data$simulation <- -9999
	}
	
	
    ##data <- merge(observation, simulation)
    ##names(data) <- c("observation", "simulation")
    
    
    data <- as.data.frame(data)
    
    ggplot(data = data, aes(x=observation, y=simulation)) +
      geom_point() +
      geom_abline(intercept=0, slope=1, col=rgb(1,0,0,.5), lwd=2)

})     
```
